$Z^\prime$ mass limits and the naturalness of supersymmetry

### 2015-05-17 paper/Referee-response.tex

The leading two-loop corrections to the effective potential are not
included in our calculation of the fine tuning.  Although the one-loop
effective potential corrections are important they are already smaller
than the RGE corrections and so we do not expect the two-loop
effective potential corrections to contribute at the same level as
the terms we include nor one which would significantly alter our
conclusions.

To address this we have added a comment specifying the precise one
loop corrections we use to the end of the paragraph below Eq.~(30).

\item {\it ``on page 15 the authors claim that for $M_X = 10^{16}$ GeV a
lower bound of 200 for the fine tuning taking $m_h = 125$~GeV. Looking
at their approximation in the calculation of $m_h$ this mass has most
likely a theory uncertainty of about 10 GeV and, thus, I do not
understand how they can make such a statement.  This also concerns
their statement on page 17.''}

The referee is correct to point out that there is a large theoretical
uncertainty on the Higgs mass and we did not mean to suggest
otherwise.  When we discuss the fine tuning dependence on the Higgs
mass we are trying to show the qualitative behavior of the tuning
increasing with the Higgs mass and how this changes depending on the
UV cutoff, using the numbers we quote to illustrate this.  We agree
that the theoretical error on the Higgs mass is rather large and that
this affects the impact the Higgs mass measurement has on the minimum
tuning that is consistent with observation.  However such discussions
raise difficult questions about how large the theoretical error is and
how to address that in the context of fine tuning.  Therefore we leave
such tricky questions to papers which are focussed on the impact on
the Higgs rather than the fine tuning from the $Z^\prime$ mass which
we are interested in here.

@@

To address the referee's remarks we have adjusted the wording of the
first comment referred to and have added a footnote pointing out that
there will be a sizable uncertainty of the Higgs mass.

\item {\it ``looking at their table II, one sees that in certain parts of
the parameter space, $A_t$ can be much larger than the corresponding soft
masses. This implies the danger of color and charge breaking minima as is
known for long time. I guess that quite some points are excluded by this.
Moreover, it seems that they allow also $|\mu|$ to be below 100 GeV which
would conflict LEP2 data. Are such points really included in their analysis?''}

For parts of the parameter space scanned the referee is correct to
point out that there is a danger of color or charge breaking minima,
and that the range of values for $\mu$ scanned over includes values
that would be in conflict with experimental limits.  In our analysis
we excluded points that may lead to such minima by removing those
which have tachyons in the spectrum, and for the MSSM we have
additionally checked that included points satisfy known conditions for
the dangerous minima to be avoided.  This however was not commented on
in our submission.  Regarding the LEP2 data and chargino mass bounds,
in Figures 2 and 4 the chargino mass limit was not imposed.  This is
because the LEP bounds are not severe enough to effect the minimum
obtainable fine tuning in the models we looked at and we wanted to
consider the impact of the chargino mass seperately in Figure 3. There
we show that much higher limits set at the LHC can have a meaningful
impact.  Nevertheless the referee is
correct in pointing out that this should be addressed in the text.

### 2015-05-17 paper/Referee-response.tex

The referee is correct to point out that there is a large theoretical
uncertainty on the Higgs mass and we did not mean to suggest
otherwise.  When we discuss the fine tuning dependence on the Higgs
mass we are trying to show the qualitative behavior of the tuning
increasing with the Higgs mass and how this changes depending on the
UV cutoff, using the numbers we quote to illustrate this.  We agree
that the theoretical error on the Higgs mass is rather large and that
this affects the impact the Higgs mass measurement has on the minimum
tuning that is consistent with observation.  However such discussions
raise difficult questions about how large the theoretical error is and
how to address that in the context of fine tuning.  Therefore we leave
such tricky questions to papers which are focussed on the impact on
the Higgs rather than the fine tuning from the $Z^\prime$ mass which
we are interested in here.

To address the referee's remarks we have adjusted the wording of the
first comment referred to and have added a footnote pointing out that
there will be a sizable uncertainty of the Higgs mass.

\item {\it ``looking at their table II, one sees that in certain parts of
the parameter space, $A_t$ can be much larger than the corresponding soft
masses. This implies the danger of color and charge breaking minima as is
known for long time. I guess that quite some points are excluded by this.
Moreover, it seems that they allow also $|\mu|$ to be below 100 GeV which
would conflict LEP2 data. Are such points really included in their analysis?''}

For parts of the parameter space scanned the referee is correct to
point out that there is a danger of color or charge breaking minima,
and that the range of values for $\mu$ scanned over includes values
that would be in conflict with experimental limits.  In our analysis
we excluded points that may lead to such minima by removing those
which have tachyons in the spectrum.  This however was not commented on
in our submission.  Regarding the LEP2 data and chargino mass bounds,
in Figures 2 and 4 the chargino mass limit was not imposed.  This is
because the LEP bounds are not severe enough to affect the minimum
obtainable fine tuning in the models we looked at and we wanted to
consider the impact of the chargino mass separately in Figure 3.  There
we show that much higher limits set at the LHC can have a meaningful
impact.  Nevertheless the referee is correct in pointing out that this should
be addressed in the text.

To address these points, a short comment has been added on page 17,
prior to Table II, in which the theoretical constraints that are applied to
the analysis are clarified.  In particular, we explain that in order to
exclude points that may have color and charge breaking minima we only allow
points that do not lead to tachyonic sparticle masses.  In addition to this,
we have added a remark on page 18 to point out that imposing the LEP bound
does not have an effect on the results, but that LHC limits can have a
significant impact.

### 2015-05-17 paper/e6ssm-tuning-paper.tex
Made changes to discussion about CCB minima and chargino limits
by deleting footnote explaining that points below the LEP limit
are included in main tuning plot, and removed comment about
additional MSSM check for CCB minima or UFB potential. Updated
referee response to match.

In this case we now consider realistic scenarios, where the
parameters that are not scanned over are set to values which keep the
associated states comfortably above their experimental limits.  So in
both the MSSM and E$_6$SSM, all other soft scalar masses are set
to $5$ TeV.  We also impose several theoretical constraints on the points
included in the analysis.  In particular, we require a valid spectrum with no
tachyonic states to exclude points which would have an unrealistic minimum,
for example due to the appearance of charge or color breaking (CCB) minima.
As a further requirement in the MSSM we exclude any points that do not satisfy
the well known CCB constraint \cite{Casas:1995pd},
\begin{equation} \label{eq:MSSMCCBconstraint}
A_t^2 < 3 \left ( \mu^2 + m_{H_u}^2 + m_{Q_3}^2 + m_{u_3}^2 \right ),
\end{equation}
or the unbounded from below (UFB) constraint,
\begin{equation} \label{eq:MSSMUFBconstraint}
m_{H_d}^2 + m_{H_u}^2 + 2 \mu^2 > 2 B \mu .
\end{equation}
We work in the third family approximation, taking the first and second
generation Yukawa couplings to be zero, and we also assume their associated
soft trilinears vanish.  Similarly we take $A_b = A_\tau = 0$ GeV.  The $U(1)$
gaugino soft mass $M_1$ was fixed to $M_1 = 300$ GeV, and we fix $M_3 = 2000$
GeV.  Additionally, in the E$_6$SSM the $U(1)_N$ gaugino soft mass $M_1'$ is
held fixed with $M_1' = M_1 = 300$ GeV, and $\mu^\prime = 5$ TeV.

@@

There is another limit which plays a similar role.  Chargino limits
directly constrain the $\mu$ parameter (or effective $\mu$ parameter
in these $U(1)$ extensions).  The LEP bound \cite{Kraan:2005vy} on
chargino masses, excluding $m_{\tilde{\chi}^\pm_1} \lesssim 104$ GeV, implies
that $|\mu|$ should only be greater than $\sim 100$ GeV, which is not
substantially larger than $M_Z$.  Consequently the bound from LEP is not
high enough to have an impact on the fine tuning obtained in the models and
parameter space regions that we have studied, as we have checked explicitly\footnote{It is for this reason that we did not require the LEP bounds be
satisfied for the points shown in Figure \ref{Fig:e6ssmvsmssm}
and in Figure \ref{Fig:othere6modelsvsmssm} below.}.  Significantly larger
lower bounds on the $\mu$ parameter, and therefore on the fine tuning, may
arise from chargino limits coming from LHC searches.  However the chargino
limits from the LHC depend on whether there are light sleptons or sneutrinos
and the mass difference between the lightest chargino and lightest neutralino.
Current limits placed by CMS and ATLAS extend up to
$m_{\tilde{\chi}^\pm_1} \approx 700 - 740$ GeV if there are light
sleptons \cite{Khachatryan:2014qwa, Aad:2014nua} with much weaker
bounds if there are no light sleptons or sneutrinos\footnote{Useful
summary plots of these limit may be found on the public pages of
ATLAS, \url{https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/CombinedSummaryPlots/SUSY/ATLAS_SUSY_EWSummary/ATLAS_SUSY_EWSummary.png}
and CMS \url{http://cms.web.cern.ch/sites/cms.web.cern.ch/files/styles/large/public/field/image/Image_03_exclusion_Combined.png?itok=8FMBpu_1}.}.

>>>>

In this case we now consider realistic scenarios, where the
parameters that are not scanned over are set to values which keep the
associated states comfortably above their experimental limits.  So in
both the MSSM and E$_6$SSM, all other soft scalar masses are set
to $5$ TeV.  We also impose several theoretical constraints on the points
included in the analysis.  In particular, we require a valid spectrum with no
tachyonic states to exclude points which would have an unrealistic minimum,
for example due to the appearance of charge or color breaking (CCB) minima.
We work in the third family approximation, taking the first and second
generation Yukawa couplings to be zero, and we also assume their associated
soft trilinears vanish.  Similarly we take $A_b = A_\tau = 0$ GeV.  The $U(1)$
gaugino soft mass $M_1$ was fixed to $M_1 = 300$ GeV, and we fix $M_3 = 2000$
GeV.  Additionally, in the E$_6$SSM the $U(1)_N$ gaugino soft mass $M_1'$ is
held fixed with $M_1' = M_1 = 300$ GeV, and $\mu^\prime = 5$ TeV.

@@

There is another limit which plays a similar role.  Chargino limits
directly constrain the $\mu$ parameter (or effective $\mu$ parameter
in these $U(1)$ extensions).  The LEP bound \cite{Kraan:2005vy} on
chargino masses, excluding $m_{\tilde{\chi}^\pm_1} \lesssim 104$ GeV, implies
that $|\mu|$ should only be greater than $\sim 100$ GeV, which is not
substantially larger than $M_Z$.  Consequently the bound from LEP is not
high enough to have an impact on the fine tuning obtained in the models and
parameter space regions that we have studied, as we have checked explicitly.
Significantly larger lower bounds on the $\mu$ parameter, and therefore on the
fine tuning, may arise from chargino limits coming from LHC searches.  However
the chargino limits from the LHC depend on whether there are light sleptons or
sneutrinos and the mass difference between the lightest chargino and lightest
neutralino.  Current limits placed by CMS and ATLAS extend up to
$m_{\tilde{\chi}^\pm_1} \approx 700 - 740$ GeV if there are light
sleptons \cite{Khachatryan:2014qwa, Aad:2014nua} with much weaker
bounds if there are no light sleptons or sneutrinos\footnote{Useful
summary plots of these limit may be found on the public pages of
ATLAS, \url{https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/CombinedSummaryPlots/SUSY/ATLAS_SUSY_EWSummary/ATLAS_SUSY_EWSummary.png}
and CMS \url{http://cms.web.cern.ch/sites/cms.web.cern.ch/files/styles/large/public/field/image/Image_03_exclusion_Combined.png?itok=8FMBpu_1}.}.
